---
title: "07_06_2021"
author: "Rebecca Netson"
date: "7/6/2021"
output: pdf_document
---

Loading in the Data for Health Variables and Lifestyle Variables
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/UML/Group Project")
library(readxl)
library(writexl)
library(gtsummary)
library(stats)
library(Hmisc)
library(ggplot2)
library(stats)
library(ggeffects)
library(ggiraphExtra)
library(dplyr)
library(lme4)
library(survey)
library(MatchIt)
library(extrafont)
library(reshape2)
library(ggpubr)
#font_import()
#loadfonts(device = "win")
df_life <- read_excel("~/UML/data_groupproject.xlsx", 
    sheet = "Lifestyle variables")
df_health <- read_excel("~/UML/data_groupproject.xlsx", 
    sheet = "Health Variables")

```
Choosing variables
```{r}
df_life_new = subset(df_life, select = -c(FDSLESS_A, ECIGW_A, SMKNOW_A, OPDFREQ_A, CIGNOW_A))
write_xlsx(df_life_new, "df_life_new.xlsx")
df_health_new = subset(df_health, select = -c(WTFA_A, CHL12M_A, CHLMED_A, HYP12M_A, HYPMED_A))
write_xlsx(df_health_new, "df_health_new.xlsx")
```

Summaries of each data frame (lifestyle and health)
```{r}
#df_life_new %>% tbl_summary()
#df_health_new %>% tbl_summary()
```

Combined data frame 
```{r}
df <- read_excel("~/UML/Group Project/final_data_group.xlsx")
df <- df[complete.cases(df),]
df =  df %>% mutate(observation = 1:n())
```

Demographic data
```{r}
df_demo = subset(df, select = c(AGEP_A, RACEALLP_A, SEX_A))
df_demo %>% describe(IQR = TRUE)
df_demo %>% tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} / {N} ({p}%)"))

library(RColorBrewer)
library(extrafont)
ggplot(df_demo,aes(x=AGEP_A, color = as.factor(RACEALLP_A))) + 
  theme_classic()+
  geom_bar(data=subset(df_demo,RACEALLP_A == "1"), alpha = 0.2) +
  geom_bar(data=subset(df_demo, RACEALLP_A == "2"), alpha = 0.2) +
  geom_bar(data=subset(df_demo, RACEALLP_A == "3"), alpha = 0.2)+
  geom_bar(data=subset(df_demo, RACEALLP_A == "4"),alpha = 0.2)+
  geom_bar(data=subset(df_demo, RACEALLP_A == "5"),alpha = 0.2)+
  geom_bar(data=subset(df_demo, RACEALLP_A == "6"),alpha = 0.2) + scale_color_brewer(name = "Ethnicity/Race Groups", labels = c("White", "Black", "Asian", "AIAN", "AIAN and other groups", "Other single or multiple race groups"), palette = "Blues") + theme(text=element_text(family = "Times New Roman", size = 10))


write_xlsx(df, "data_group_project_completecases.xlsx")
```

Re-coding into binary variables
(yes = 1) (no = 0)
```{r}
df$SMKEV_A <- ifelse(df$SMKEV_A == 1, 1, 0) #smoked at least 100 cigarettes
df$PREDIB_A <- ifelse(df$PREDIB_A == 1, 1, 0) #pre diabetes
df$DIBEV_A <- ifelse(df$DIBEV_A == 1, 1, 0) #diabetes
df$ASEV_A <- ifelse(df$ASEV_A == 1, 1, 0) #asthma 
df$CHDEV_A <- ifelse(df$CHDEV_A == 1, 1, 0) #coronary heart disease
df$SEX_A <- ifelse(df$SEX_A == 1, 1, 0)#gender where females = 0
```

Propensity Matching
```{r}
#df$BMICAT_A <- ifelse(df$BMICAT_A == 3|df$BMICAT_A == 4, 1, 0)
#df.out <- matchit(BMICAT_A ~ AGEP_A + RACEALLP_A + SEX_A, data = df, method = "nearest", distance = "glm")
#matched.data = match.data(df.out)
#sum(matched.data$BMICAT_A == 1) #463 people that are overweight or obese
#sum(matched.data$BMICAT_A == 0) #463 people who are underweight or normal

#plot(summary(df.out))

#low_bmi_df = matched.data[matched.data$BMICAT_A == 0, ]
#high_bmi_df = matched.data[matched.data$BMICAT_A == 1, ]
#low_bmi_df$SEX_A = as.factor(low_bmi_df$SEX_A)
#high_bmi_df$SEX_A = as.factor(high_bmi_df$SEX_A)
#low_bmi_df = subset(low_bmi_df, select = -c(observation, distance, weights, subclass))
#high_bmi_df = subset(high_bmi_df, select = -c(observation, distance, weights, subclass))
#low_bmi_df %>% tbl_summary(by = SEX_A)
#high_bmi_df %>% tbl_summary(by = SEX_A)
#head(low_bmi_df)
#t.test(low_bmi_df$CHDEV_A, high_bmi_df$CHDEV_A)


#ggplot(matched.data, aes(x = CHDEV_A, y= AGEP_A, color = as.factor(BMICAT_A))) + geom_bar(stat="identity")
#mod1 <- glm(CHDEV_A ~ AGEP_A + SEX_A +  PHSTAT_A, data = matched.data, family = binomial)
#summary(mod1)
```


```{r}
#making into factors
df$SEX_A <- as.factor(df$SEX_A)
df$BMICAT_A <- as.numeric(df$BMICAT_A)
contrasts(df$SEX_A)
df$SMKEV_A = as.factor(df$SMKEV_A)
df$RACEALLP_A = as.factor(df$RACEALLP_A)
df$PHSTAT_A = as.numeric(df$PHSTAT_A)
df$AGEP_A = as.numeric(df$AGEP_A)
df$BMICAT_A = as.factor(df$BMICAT_A)
df$SMKCIGST_A = as.numeric(df$SMKCIGST_A)

#making age groups 
df$AgeGroup = vector("double", length(nrow(df)))
for(i in 1:nrow(df)){
  if(df$AGEP_A[i] > 17.9 & df$AGEP_A[i] <= 29.2){
    df$AgeGroup[i] = 1
  } else if(df$AGEP_A[i] > 29.2 & df$AGEP_A[i] <= 40.3){
    df$AgeGroup[i] = 2
  } else if(df$AGEP_A[i] > 40.3 & df$AGEP_A[i] <= 51.5 ){
    df$AgeGroup[i] = 3
  } else if(df$AGEP_A[i] > 51.5 & df$AGEP_A[i] <= 62.7){
    df$AgeGroup[i] = 4  
  } else if(df$AGEP_A[i] > 62.7 & df$AGEP_A[i] <= 73.8){
    df$AgeGroup[i] = 5
  } else{
    df$AgeGroup[i] = 6 #Age of participant is 73.8 < Age < 85.1
  }
}
df$AgeGroup = as.factor(df$AgeGroup)



```

Simple Logistic Regression
```{r}
#mod1 <- glm(CHDEV_A ~ AGEP_A + SEX_A +  PHSTAT_A, data = df, family = binomial)
#summary(mod1)
#summary(mod1)$coef
#probabilities <- mod1 %>% predict(df, type = "response")
#head(probabilities)
#newdata <- data.frame(AGEP_A = seq(min(df$AGEP_A), max(df$AGEP_A), len = 1352), PHSTAT_A = seq(min(df$PHSTAT_A), max(df$PHSTAT_A), len = 1352), RACEALLP_A = seq(min(df$RACEALLP_A), max(df$RACEALLP_A), len = 1352))
#newdata$ASEV_A = predict(mod1, newdata=newdata, type = "response")

#plot(ASEV_A ~ AGEP_A, newdata)                                                                                                     
                                                                                                                                                                  
#probabilities <- mod1 %>% predict(newdata, type = "response")
#plot(probabilities)
#predicted.classes <- ifelse(probabilities > 0.1, "Yes", "No")
#predicted.classes
#p = df %>% 
    #ggplot(aes(AGEP_A, probabilities)) +
    #geom_point(alpha = 0.2) + ylim(0, 0.30) + 
    #geom_smooth(method = "glm", method.args = list(family = "binomial"))
#p    

#q = ggPredict(mod1, se = TRUE, interactive = FALSE, digits = 2)
#q
#mod2 <- glm(ASEV_A ~ AGEP_A, data = df, family = binomial)
#newdata2 <- data.frame(AGEP_A = c(20, 80))
#probabilities2 <- mod2 %>% predict(newdata2, type = "response")
#predicted.classes2 <- ifelse(probabilities2 > 0.1, "pos", "neg")
#predicted.classes2
#df %>%
  #mutate(prob = ifelse(ASEV_A == 1, 1, 0)) %>%
  #ggplot(aes(AGEP_A, prob)) +
  #geom_point(alpha = 0.2) +
  #geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  #labs(
    #title = "Logistic Regression Model", 
    #x = "Plasma Glucose Concentration",
    #y = "Probability of being diabete-pos"
    #)


```


Comparing groups 
```{r}
#split by Gender (female = 2, male = 1)
df_females <-  df[df$SEX_A == 0, ]
df_males <- df[df$SEX_A == 1, ]
#df_males %>% describe
#by_gender = df %>% tbl_summary(by  = SEX_A) %>% add_p()
#by_gender = as_tibble(by_gender)
t.test(df_females$SMKCIGST_A, df_males$SMKCIGST_A)
#write_xlsx(by_gender, "by_gender.xlsx")

#split by race groups 
df_white <- df[df$RACEALLP_A == 1, ]
df_black <- df[df$RACEALLP_A == 2, ]
df_asian <- df[df$RACEALLP_A == 3, ]
df_aian <- df[df$RACEALLP_A == 4, ]
df_aian_other <- df[df$RACEALLP_A == 5, ]
df_females

df_obese = df[df$BMICAT_A == 4|df$BMICAT_A == 3, ]
df_normal = df[df$BMICAT_A == 2, ]
df_obese$CHDEV_A = as.factor(df_obese$CHDEV_A)
#ggplot(df_obese, aes(x = AGEP_A, y = PHSTAT_A, color = CHDEV_A)) + geom_point + geom_line(m)
mod <- glm(CHDEV_A ~ AGEP_A + SEX_A, data = df_obese, family = binomial)
summary(mod)$coef
probabilities <- mod %>% predict(df_obese, type = "response")
df_obese$prob_of_CHDEV_A = probabilities


mod1 <- glm(CHDEV_A ~ AGEP_A + SEX_A, data = df_normal, family = binomial)
df_normal$prob_of_CHDEV_A = mod1 %>% predict(df_normal, type = "response")
summary(mod1)
probabilities1 <- mod1 %>% predict(df, type = "response")
df1 = subset(df_obese, select = c(AGEP_A, SEX_A, prob_of_CHDEV_A))
df2 = subset(df_normal, select = c(AGEP_A, SEX_A, prob_of_CHDEV_A))
names(df1) <- c("Age", "Sex", "Prob")
names(df2) <- c("Age", "Sex", "Prob")
dat <- rbind(df1, df2)
dat$Group <- rep(factor(1:2), times  = c(nrow(df1),nrow(df2)))
dat$GroupName <- vector("character",length = nrow(dat))
dat$GroupName <- ifelse(dat$Group == 1, "Obese or Overweight", "Healthy Weight")
dat
t.test(df_obese$AGEP_A, df_normal$AGEP_A, method = "pearson")
q <- ggplot(data = dat, aes(x = Age, y = Prob, color = as.factor(Sex)))+geom_point(aes(shape = GroupName), size = 2)+theme(text=element_text(family = "Times New Roman", size = 10))
q
q + scale_color_discrete(name = "Sex",
                         labels = c("Females", "Males")) + labs(
    title = "Logistic Regression Model: Cardiovascular Health ~ Age by Sex with Groups", 
    x = "Age",
    y = "Probability"
    )  + theme(legend.position = "top") + scale_size_manual(values = c(2,3))



#tbl_summary(df_obese)
#tbl_summary(df_normal)

mod_asthma_normal <-glm(ASEV_A ~ AGEP_A + SEX_A, data = df_normal, family = binomial)
df_normal$prob_of_asthma = mod_asthma_normal %>% predict(df_normal, type = "response")



mod_asthma_obese <- glm(ASEV_A ~ AGEP_A + SEX_A, data = df_obese, family = binomial)
df_obese$prob_of_asthma = mod_asthma_obese %>% predict(df_obese, type = "response")

df_n = subset(df_normal, select = c(AGEP_A, SEX_A, prob_of_asthma))
df_o = subset(df_obese, select = c(AGEP_A, SEX_A, prob_of_asthma))

dat_asthma <- rbind(df_n, df_o)
dat_asthma$Group <- rep(factor(1:2), times  = c(nrow(df_n),nrow(df_o)))
dat_asthma$GroupName <- vector("character",length = nrow(dat_asthma))
dat_asthma$GroupName <- ifelse(dat_asthma$Group == 1, "Healthy Weight", "Overweight or Obese")
dat_asthma

#By Race:
by_race = df %>% tbl_summary(by = RACEALLP_A) 
by_race = as_tibble(by_race)
#write_xlsx(by_race, "by_race.xlsx")



library(RColorBrewer)

r <- ggplot(data = dat_asthma, aes(x = AGEP_A, y = prob_of_asthma, color = as.factor(SEX_A)))+geom_point(aes(shape = GroupName), size = 2)+theme(text=element_text(family = "Times New Roman", size = 10))
r
r = r + scale_color_brewer(name = "Sex",
                         labels = c("Females", "Males"), palette = "Dark2") + xlab("Age") + ylab("Probability") +  theme(legend.position = "top") + ggtitle("Logistic Regression Model: Asthma ~ Age by Sex with Groups")+scale_x_continuous(breaks = round(seq(min(dat_asthma$AGEP_A), max(dat_asthma$AGEP_A), by = 5),85))



df$BMICAT_A = as.factor(df$BMICAT_A)
df$SMKCIGST_A = as.numeric(df$SMKCIGST_A)

ggplot(df_females, aes(x = ))
```

General health predictor model
```{r}
#checking normality of predictor variables
shapiro.test(df$PHSTAT_A)
df$SMKEV_A = as.numeric(df$SMKEV_A)
gen_model = lm(PHSTAT_A ~ AGEP_A + FDSBALANCE_A + UPPSLFCR_A + SMKEV_A + BMICAT_A, data = df)
summary(gen_model)

```

Health Characteristics Graph

```{r}
df$BMICAT_A = as.numeric(df$BMICAT_A)
dfem1 <- df_females %>%
  group_by(BMICAT_A, PREDIB_A) %>%
  summarise(count = n())
dfmal1 <- df_males %>%
  group_by(BMICAT_A, PREDIB_A) %>%
  summarise(count = n())



dfem1$PREDIB_A = as.factor(dfem1$PREDIB_A)
dfmal1
ggplot(dfem1, aes(x = PREDIB_A, y = count)) + geom_bar() 
  
dfem1
p= ggplot(dfem1, aes(factor(BMICAT_A), count, fill = PREDIB_A)) + geom_col(position = 'dodge') + theme_classic()+ theme(text=element_text(family = "Times New Roman", size = 10)) + labs(fill = "Prediabetes") + xlab("Body Mass Index") + ylab("Count by Group") + scale_fill_discrete(labels = c("Answered Yes", "Answered No"))+ scale_x_discrete(labels = c("Overweight", "Healthy Weight", "Overweight", "Obese")) + ggtitle("Female Health: Body Mass Index ~ Prediabetes Count")

dfmal1 <- as.data.frame(dfmal1)
dfmal1$PREDIB_A = as.factor(dfmal1$PREDIB_A)
dfmal1
p= ggplot(dfmal1, aes(factor(BMICAT_A), count, fill = PREDIB_A)) + geom_col(position = 'dodge') + theme_classic()+ theme(text=element_text(family = "Times New Roman", size = 10)) + labs(fill = "Prediabetes") + xlab("Body Mass Index") + ylab("Count by Group") + scale_fill_discrete(labels = c("Answered Yes", "Answered No"))+ scale_x_discrete(labels = c("Overweight", "Healthy Weight", "Overweight", "Obese")) + ggtitle("Male Health: Body Mass Index ~ Prediabetes Count")
```
Life characteristics Graph
```{r}
l.fem <- df_females %>%
  group_by(FDSBALANCE_A, SMKCIGST_A) %>%
  summarise(count = n())
l.fem$FDSBALANCE_A = as.factor(l.fem$FDSBALANCE_A)
l.fem$SMKCIGST_A = as.numeric(l.fem$SMKCIGST_A)
l.fem
ggplot(l.fem, aes(factor(SMKCIGST_A), count, fill = FDSBALANCE_A)) + geom_col(position = 'dodge') + theme_classic()+ theme(text=element_text(family = "Times New Roman", size = 10)) + labs(fill = "Ability to eat balanced meals") + xlab("Smoking Status") + ylab("Count by Group") + scale_fill_discrete(labels = c("Often True", "Sometimes True", "Never True"))+ scale_x_discrete(labels = c("Every day smoker", "Some day smoker", "Former smoker", "Never smoker")) + ggtitle("Female Lifestyle: Smoking status ~ Ability to eat balanced meals by Count")

l.male <- df_males %>%
  group_by(FDSBALANCE_A, SMKCIGST_A) %>%
  summarise(count = n())
l.male$FDSBALANCE_A = as.factor(l.male$FDSBALANCE_A)
l.male$SMKCIGST_A = as.numeric(l.male$SMKCIGST_A)
l.male
ggplot(l.male, aes(factor(SMKCIGST_A), count, fill = FDSBALANCE_A)) + geom_col(position = 'dodge') + theme_classic()+ theme(text=element_text(family = "Times New Roman", size = 10)) + labs(fill = "Ability to eat balanced meals") + xlab("Smoking Status") + ylab("Count by Group") + scale_fill_discrete(labels = c("Often True", "Sometimes True", "Never True"))+ scale_x_discrete(labels = c("Every day smoker", "Some day smoker", "Former smoker", "Never smoker")) + ggtitle("Male Lifestyle: Smoking status ~ Ability to eat balanced meals by Count")



```
by age group
```{r}
df$SEX_A = as.factor(df$SEX_A)
df$BMICAT_A = as.numeric(df$BMICAT_A)
by_age = df %>%
  group_by(AgeGroup, SEX_A) %>%
  summarise(
  count = n(),
  mean_weight = mean(BMICAT_A, na.rm = TRUE),
  mean_smoking = mean(SMKCIGST_A, na.rm = TRUE),
  mean_anxiety = mean(ANXFREQ_A, na.rm = TRUE)
)
by_age = as.data.frame(by_age)

library(multcomp)
library(stats)
df$AgeGroup = as.factor(df$AgeGroup)
age.aov <- aov(ANXFREQ_A~ AgeGroup, data = df)
TukeyHSD(age.aov)
age.aov

summary(glht(age.aov, linfct=mcp(AgeGroup = "Tukey")))
```


